### 2020-03-06 submission/manuscript.Rmd
Minor edits

In our early Code Clubs we noticed that if the presenter did not clearly state their goals for the session, it would often led to frustration for the presenter and participants. If the presenter shared their own code, did they want us to focus on their coding style or do they want help with incorporating a new package into their workflow? Participants will always notice or ask about things that are not the focus of the exercise. Be specific about the goal and bring them back to the task if they are distracted. Where possible, creating a simplified scenario (i.e. minimal, reproducible example [@StackOverflow_MREE]) can be helpful to remove distractions. The presenter should make sure that the simplified example works and behaves the way they expect before coming to the Code Club. Beyond the content, participants should know what they can be expected to have completed by the end of the session.

@@

Our Code Club needs to fit within an hour time slot. The presenter should anticipate enough time for an introduction and brief instruction, time for participants to engage the material, and time for everyone to report back. A typical schedule for Code Review is 10 minutes of introduction and instruction, 30 minutes of paired programming, 5 minutes to get groups to wrap up, and 10 minutes to report back to the group. Remember that a task that you can accomplish in 10 minutes may take a more junior member of your group 30 minutes. The goal of the Code Club should be to help that junior member to learn the topic and to reinforce the topic for the more senior person and help them see the concept in a new context. One possibility would be to create "stretch goals" for those that complete the task more quickly than everyone else. We once had a presenter try to teach the group some basic Julia syntax. By the time people installed the interpreter, the time was up. Alternatively, a session where participants were asked to make code DRY gave them multiple stopping points where they could feel successful (e.g. commenting code, creating function, implementing function, refactoring function).

@@

Similar to a Journal Club, the presenter should give participants a few days to prepare for the Code Club. Be sure to tell the participants what you will be covering. Provide the them with instructions on how to install dependencies, download data, and get the initial code. Make sure that the materials work together as intended. Considering the compressed schedule, asking participants to download materials beforehand is helpful. We have found that using separate GitHub repositories for each session can be a helpful way to get information to participants. Of course, using git and GitHub to engage in collaborative coding should probably be its own Code Club before using the tools more broadly with the group. An easier entry point is to have groups post their code to a Slack channel that we use to announce materials for our meetings.

@@

It is the rare scientist that can fully participate in a Journal Club discussion about a paper that they have not read. Coming to Code Club without installing a dependency is like going to Journal Club and asking a simple question about the paper. It shows a lack of respect and preparation. Following Rule 4, review the code in advance, download the datasets, install the necessary packages, and perhaps read up on the topic. If your Code Club is reading a paper or chapters in a data science book be sure to read them before the session (e.g. [@Wickham_2016; @Schloss_2018; @Noble2009]). Consider how you could use the material for your own work.

@@

A significant challenge to presenting at Code Club is to pick topics and activities that appeal to a critical mass of the participants. For example, if there has recently been a large turn over in group membership, it can be useful to go back to the basics and cover commonly used tools. We have identified several strategies to overcome the challenges of having varying skill levels. As an alternative to letting participants form their own pairs, the presenter can be purposeful in how they form pairs. Depending on the goals of the session, forming pairs between participants that have similar or different skills can be effective. If pairs are formed where partners have similar skill levels, then the presenter needs to design activities that are appropriate to different levels of participants. We have found that a good activity for beginners is to have them comment code, which forces them to dissect a code chunk and understand what each line does. An advantage of forming pairs with disparate skill levels is that all of the pairs can be given the same task leading to a diversity of methods to achieve the same result. This approach to pairing also helps graft new members into the group that have emerging programming skills. Regardless of how partners are selected, asking the pairs to identify a navigator and driver is helpful [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type. Midway through the activity the presenter can tell the partners to switch roles. Intentionally forming pairs can also be a mechanism to engineer interactions by breaking up cliques, avoiding potentially toxic combinations of members, or pairing reliable role models with new members of the group.

@@

Having an open laptop on the table and permission to use it can feel like an invitation to get distracted by other work, emails, and browsing the internet. Fight that urge and focus on the presenter's goals. Allow your partner to contribute. Speak up for yourself and force your partner to let you contribute. Even if the material seems too advanced for you, programming languages like R and Python are generally expressive and should allow you to engage with the logic, even if the syntax is too hard. Frankly, understanding the logic of when to use one modeling approach over another is more important than how to use the modeling approach. If you understand the "why", the "how" will generally quickly follow. If you are a more experienced participant, give your feedback and coding suggestions at a level that seeks to inform everyone. Regardless of your skill level, your partner and the presenter are putting themselves at risk by revealing what they do or do not know. Encourage them and show your gratitude for helping you learn something new.

@@

Part of the humility required to participate in a Code Club is both acknowledging your incomplete training and that learning something new is possible. Assume that the presenter has followed the other rules and that they think you should be using what they presented. After the Code Club, try to incorporate that material into your new code or by refactoring your old code. By practicing the material in a different context you will learn the material better. You may identify concepts that you are unsure of, which gives you another opportunity to ask the presenter for help.

@@

The selection of content and structure for each Code Club is determined by the presenter making the process democratic and distributed. If someone thinks something is worth learning and wants to teach it, they have that power. If they want to experiment with a different format, they are free to try it out. Members of the Code Club need to feel like they have the power to shape the direction of the group. If members are following these Rules, they will naturally reflect on the skills and interests of the other members in the group. For example, there is always turn over in a research group and thus it is important to make sure that there are opportunities to go back to things people consider "simple" to teach the content to new people or provide a refresher to people that perhaps have been at the lab bench more than at the computer recently. The group should also feel free to experiment with the format. If we try a different format, then it is useful to give everyone an opportunity to debrief on what they did or did not like with the new approach.

>>>>

In our early Code Clubs we noticed that if the presenter did not clearly state their goals for the session, it would often lead to frustration for the presenter and participants. If the presenter shared their own code, did they want us to focus on their coding style or did they want help with incorporating a new package into their workflow? Participants will always notice or ask about things that are not the focus of the exercise. Be specific about the goals and bring them back to the task if they are distracted. Where possible, creating a simplified scenario (i.e. minimal, reproducible example [@StackOverflow_MREE]) can be helpful to remove distractions. The presenter should make sure that the simplified example works and behaves the way they expect before coming to the Code Club. Beyond the content, participants should know what they can be expected to have completed by the end of the session.

@@

Our Code Club needs to fit within an hour time slot. The presenter should anticipate enough time for an introduction and brief instruction, time for participants to engage the material, and time for everyone to report back. A typical schedule for Code Review is 10 minutes of introduction and instruction, 30 minutes of paired programming, 5 minutes to get groups to wrap up, and 10 minutes to report back to the group. Remember that a task that you can accomplish in 10 minutes may take a more junior member of your group 30 minutes. The goal of the Code Club should be to help that junior member to learn the topic and to reinforce the topic for the more senior person and help them see the concept in a new context. One possibility would be to create "stretch goals" for those that complete the task more quickly than everyone else. We once had a presenter try to teach the group some basic Julia syntax. By the time people installed the interpreter, the time was up. Alternatively, a session where participants were asked to make code DRY gave them multiple stopping points where they could feel successful (e.g. documenting code, writing a function signature, implementing a function, refactoring code).

@@

Similar to a Journal Club, the presenter should give participants a few days to prepare for the Code Club. Be sure to tell the participants what you will be covering. Provide them with instructions on how to install dependencies, download data, and get the initial code. Make sure that the materials work together as intended. Considering the compressed schedule, asking participants to download materials beforehand is helpful. We have found that using separate GitHub repositories for each session can be a helpful way to get information to participants. Of course, using git and GitHub to engage in collaborative coding should probably be its own Code Club before using the tools more broadly with the group. An easier entry point is to have groups post their code to a Slack channel that we use to announce materials for our meetings.

@@

It is the rare scientist that can fully participate in a Journal Club discussion about a paper that they have not read. Coming to Code Club without installing a dependency is like going to Journal Club and asking a simple question about the paper. It shows a lack of respect and preparation. Following Rule 4, review the code in advance, download the datasets, install the necessary packages, and perhaps read up on the topic. If your Code Club is reading a paper or chapters in a data science book, be sure to read them before the session (e.g. [@Wickham_2016; @Schloss_2018; @Noble2009]). Consider how you could use the material for your own work.

@@

A significant challenge to presenting at Code Club is to pick topics and activities that appeal to a critical mass of the participants. For example, if there has recently been a large turn over in group membership, it can be useful to go back to the basics and cover commonly used tools. We have identified several strategies to overcome the challenges of having varying skill levels. As an alternative to letting participants form their own pairs, the presenter can be purposeful in how they form pairs. Depending on the goals of the session, forming pairs between participants that have similar or different skills can be effective. If pairs are formed where partners have similar skill levels, then the presenter needs to design activities that are appropriate to different levels of participants. We have found that a good activity for beginners is to have them comment code, which forces them to dissect a code chunk and understand what each line does. An advantage of forming pairs with disparate skill levels is that all of the pairs can be given the same task leading to a diversity of methods to achieve the same result. This approach to pairing also helps graft new members that have emerging programming skills into the group. Regardless of how partners are selected, asking the pairs to identify a navigator and driver is helpful [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type. Midway through the activity, the partners can switch roles. Intentionally forming pairs can also be a mechanism to engineer interactions by breaking up cliques, avoiding potentially toxic combinations of members, or pairing reliable role models with new members of the group.

@@

Having an open laptop on the table and permission to use it can feel like an invitation to get distracted by other work, emails, and browsing the internet. Fight that urge and focus on the presenter's goals. Allow your partner to contribute. Speak up for yourself and force your partner to let you contribute. Even if the material seems too advanced for you, programming languages like R and Python are generally expressive and should allow you to engage with the logic, even if the syntax is too hard. Frankly, understanding the logic of _when_ to use one modeling approach over another is more important than knowing _how_ to use the modeling approach. If you understand the "why", the "how" will generally quickly follow. If you are a more experienced programmer, give your feedback and coding suggestions at a level that seeks to inform everyone. Regardless of your skill level, your partner and the presenter are putting themselves in a vulnerable position by revealing what they do or do not know. Encourage them and show your gratitude for helping you learn something new.

@@

Part of the humility required to participate in a Code Club is both acknowledging your incomplete training and that learning something new is possible. Assume that the presenter has followed the other rules and that they think you should be using what they presented. After the Code Club, try to incorporate that material into your new code or by refactoring your old code. By practicing the material in a different context, you will learn the material better. You may identify concepts that you are unsure of, which gives you another opportunity to ask the presenter for help.

@@

The selection of content and structure for each Code Club is determined by the presenter, making the process democratic and distributed. If someone thinks something is worth learning and wants to teach it, they have that power. If they want to experiment with a different format, they are free to try it out. Members of the Code Club need to feel like they have the power to shape the direction of the group. If members are following these Rules, they will naturally reflect on the skills and interests of the other members in the group. For example, there is always turn over in a research group and thus it is important to make sure that there are opportunities to go back to things people consider "simple" to teach the content to new people or provide a refresher to people that perhaps have been at the lab bench more than at the computer recently. The group should also feel free to experiment with the format. If we try a different format, then it is useful to give everyone an opportunity to debrief on what they did or did not like with the new approach.

### 2020-03-06 submission/manuscript.tex
Minor edits

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it would often led to frustration for
the presenter and participants. If the presenter shared their own code,
did they want us to focus on their coding style or do they want help
with incorporating a new package into their workflow? Participants will
always notice or ask about things that are not the focus of the
exercise. Be specific about the goal and bring them back to the task if
they are distracted. Where possible, creating a simplified scenario
(i.e.~minimal, reproducible example {[}4{]}) can be helpful to remove
distractions. The presenter should make sure that the simplified example
works and behaves the way they expect before coming to the Code Club.
Beyond the content, participants should know what they can be expected
to have completed by the end of the session.

@@

Our Code Club needs to fit within an hour time slot. The presenter
should anticipate enough time for an introduction and brief instruction,
time for participants to engage the material, and time for everyone to
report back. A typical schedule for Code Review is 10 minutes of
introduction and instruction, 30 minutes of paired programming, 5
minutes to get groups to wrap up, and 10 minutes to report back to the
group. Remember that a task that you can accomplish in 10 minutes may
take a more junior member of your group 30 minutes. The goal of the Code
Club should be to help that junior member to learn the topic and to
reinforce the topic for the more senior person and help them see the
concept in a new context. One possibility would be to create ``stretch
goals'' for those that complete the task more quickly than everyone
else. We once had a presenter try to teach the group some basic Julia
syntax. By the time people installed the interpreter, the time was up.
Alternatively, a session where participants were asked to make code DRY
gave them multiple stopping points where they could feel successful
(e.g.~commenting code, creating function, implementing function,
refactoring function).

@@

It is the rare scientist that can fully participate in a Journal Club
discussion about a paper that they have not read. Coming to Code Club
without installing a dependency is like going to Journal Club and asking
a simple question about the paper. It shows a lack of respect and
preparation. Following Rule 4, review the code in advance, download the
datasets, install the necessary packages, and perhaps read up on the
topic. If your Code Club is reading a paper or chapters in a data
science book be sure to read them before the session (e.g.~{[}5--7{]}).
Consider how you could use the material for your own work.

@@

A significant challenge to presenting at Code Club is to pick topics and
activities that appeal to a critical mass of the participants. For
example, if there has recently been a large turn over in group
membership, it can be useful to go back to the basics and cover commonly
used tools. We have identified several strategies to overcome the
challenges of having varying skill levels. As an alternative to letting
participants form their own pairs, the presenter can be purposeful in
how they form pairs. Depending on the goals of the session, forming
pairs between participants that have similar or different skills can be
effective. If pairs are formed where partners have similar skill levels,
then the presenter needs to design activities that are appropriate to
different levels of participants. We have found that a good activity for
beginners is to have them comment code, which forces them to dissect a
code chunk and understand what each line does. An advantage of forming
pairs with disparate skill levels is that all of the pairs can be given
the same task leading to a diversity of methods to achieve the same
result. This approach to pairing also helps graft new members into the
group that have emerging programming skills. Regardless of how partners
are selected, asking the pairs to identify a navigator and driver is
helpful {[}8{]}. The driver types at the computer while the navigator
tells them what to type. Midway through the activity the presenter can
tell the partners to switch roles. Intentionally forming pairs can also
be a mechanism to engineer interactions by breaking up cliques, avoiding
potentially toxic combinations of members, or pairing reliable role
models with new members of the group.

@@

Having an open laptop on the table and permission to use it can feel
like an invitation to get distracted by other work, emails, and browsing
the internet. Fight that urge and focus on the presenter's goals. Allow
your partner to contribute. Speak up for yourself and force your partner
to let you contribute. Even if the material seems too advanced for you,
programming languages like R and Python are generally expressive and
should allow you to engage with the logic, even if the syntax is too
hard. Frankly, understanding the logic of when to use one modeling
approach over another is more important than how to use the modeling
approach. If you understand the ``why'', the ``how'' will generally
quickly follow. If you are a more experienced participant, give your
feedback and coding suggestions at a level that seeks to inform
everyone. Regardless of your skill level, your partner and the presenter
are putting themselves at risk by revealing what they do or do not know.
Encourage them and show your gratitude for helping you learn something
new.

@@

The selection of content and structure for each Code Club is determined
by the presenter making the process democratic and distributed. If
someone thinks something is worth learning and wants to teach it, they
have that power. If they want to experiment with a different format,
they are free to try it out. Members of the Code Club need to feel like
they have the power to shape the direction of the group. If members are
following these Rules, they will naturally reflect on the skills and
interests of the other members in the group. For example, there is
always turn over in a research group and thus it is important to make
sure that there are opportunities to go back to things people consider
``simple'' to teach the content to new people or provide a refresher to
people that perhaps have been at the lab bench more than at the computer
recently. The group should also feel free to experiment with the format.
If we try a different format, then it is useful to give everyone an
opportunity to debrief on what they did or did not like with the new
approach.

>>>>

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it would often lead to frustration
for the presenter and participants. If the presenter shared their own
code, did they want us to focus on their coding style or did they want
help with incorporating a new package into their workflow? Participants
will always notice or ask about things that are not the focus of the
exercise. Be specific about the goals and bring them back to the task if
they are distracted. Where possible, creating a simplified scenario
(i.e.~minimal, reproducible example {[}4{]}) can be helpful to remove
distractions. The presenter should make sure that the simplified example
works and behaves the way they expect before coming to the Code Club.
Beyond the content, participants should know what they can be expected
to have completed by the end of the session.

@@

Our Code Club needs to fit within an hour time slot. The presenter
should anticipate enough time for an introduction and brief instruction,
time for participants to engage the material, and time for everyone to
report back. A typical schedule for Code Review is 10 minutes of
introduction and instruction, 30 minutes of paired programming, 5
minutes to get groups to wrap up, and 10 minutes to report back to the
group. Remember that a task that you can accomplish in 10 minutes may
take a more junior member of your group 30 minutes. The goal of the Code
Club should be to help that junior member to learn the topic and to
reinforce the topic for the more senior person and help them see the
concept in a new context. One possibility would be to create ``stretch
goals'' for those that complete the task more quickly than everyone
else. We once had a presenter try to teach the group some basic Julia
syntax. By the time people installed the interpreter, the time was up.
Alternatively, a session where participants were asked to make code DRY
gave them multiple stopping points where they could feel successful
(e.g.~documenting code, writing a function signature, implementing a
function, refactoring code).

@@

It is the rare scientist that can fully participate in a Journal Club
discussion about a paper that they have not read. Coming to Code Club
without installing a dependency is like going to Journal Club and asking
a simple question about the paper. It shows a lack of respect and
preparation. Following Rule 4, review the code in advance, download the
datasets, install the necessary packages, and perhaps read up on the
topic. If your Code Club is reading a paper or chapters in a data
science book, be sure to read them before the session (e.g.~{[}5--7{]}).
Consider how you could use the material for your own work.

@@

A significant challenge to presenting at Code Club is to pick topics and
activities that appeal to a critical mass of the participants. For
example, if there has recently been a large turn over in group
membership, it can be useful to go back to the basics and cover commonly
used tools. We have identified several strategies to overcome the
challenges of having varying skill levels. As an alternative to letting
participants form their own pairs, the presenter can be purposeful in
how they form pairs. Depending on the goals of the session, forming
pairs between participants that have similar or different skills can be
effective. If pairs are formed where partners have similar skill levels,
then the presenter needs to design activities that are appropriate to
different levels of participants. We have found that a good activity for
beginners is to have them comment code, which forces them to dissect a
code chunk and understand what each line does. An advantage of forming
pairs with disparate skill levels is that all of the pairs can be given
the same task leading to a diversity of methods to achieve the same
result. This approach to pairing also helps graft new members that have
emerging programming skills into the group. Regardless of how partners
are selected, asking the pairs to identify a navigator and driver is
helpful {[}8{]}. The driver types at the computer while the navigator
tells them what to type. Midway through the activity, the partners can
switch roles. Intentionally forming pairs can also be a mechanism to
engineer interactions by breaking up cliques, avoiding potentially toxic
combinations of members, or pairing reliable role models with new
members of the group.

@@

Having an open laptop on the table and permission to use it can feel
like an invitation to get distracted by other work, emails, and browsing
the internet. Fight that urge and focus on the presenter's goals. Allow
your partner to contribute. Speak up for yourself and force your partner
to let you contribute. Even if the material seems too advanced for you,
programming languages like R and Python are generally expressive and
should allow you to engage with the logic, even if the syntax is too
hard. Frankly, understanding the logic of \emph{when} to use one
modeling approach over another is more important than knowing \emph{how}
to use the modeling approach. If you understand the ``why'', the ``how''
will generally quickly follow. If you are a more experienced programmer,
give your feedback and coding suggestions at a level that seeks to
inform everyone. Regardless of your skill level, your partner and the
presenter are putting themselves in a vulnerable position by revealing
what they do or do not know. Encourage them and show your gratitude for
helping you learn something new.

@@

The selection of content and structure for each Code Club is determined
by the presenter, making the process democratic and distributed. If
someone thinks something is worth learning and wants to teach it, they
have that power. If they want to experiment with a different format,
they are free to try it out. Members of the Code Club need to feel like
they have the power to shape the direction of the group. If members are
following these Rules, they will naturally reflect on the skills and
interests of the other members in the group. For example, there is
always turn over in a research group and thus it is important to make
sure that there are opportunities to go back to things people consider
``simple'' to teach the content to new people or provide a refresher to
people that perhaps have been at the lab bench more than at the computer
recently. The group should also feel free to experiment with the format.
If we try a different format, then it is useful to give everyone an
opportunity to debrief on what they did or did not like with the new
approach.
